---
output: html_document
---
Read in and create data
```{r}
a <- read.table("http://www.stat.ucla.edu/~nchristo/statistics_c173_c273/jura.txt", header=TRUE)
y <- a$Pb

x1 <- a$Cd
x2 <- a$Co
x3 <- a$Cr
x4 <- a$Cu
x5 <- a$Ni
x6 <- a$Zn

ones <- rep(1, nrow(a))
X <- as.matrix(cbind(ones,x1, x2, x3, x4, x5, x6))
X1 <- as.matrix(cbind(ones,x1, x2, x3, x4))
X2 <- as.matrix(cbind(x5, x6))
```

### Question i
Part 1:
```{r}
#H1 Matrix
H1 <- X1 %*% solve(t(X1) %*% X1) %*% t(X1)
#H matrix
H <- X %*% solve(t(X) %*% X) %*% t(X)

#H-H1 Matrix
mat <- round((H - H1) %*% (H - H1))

identical(round((H - H1)), mat)
```

The matrix (H-H1)(H-H1) is the same as the matrix (H-H1) which means that (H-H1) must be idempotent.

Part 2:

```{r}
#Beta hat using full regression
beta_hat <- solve(t(X) %*% X) %*% t(X) %*% y

H1 <- X1 %*% solve(t(X1) %*% X1) %*% t(X1)
X22 <- (diag(nrow(a)) - H1) %*% X2
y11 <- (diag(nrow(a)) - H1) %*% y

#Beta hat 1 from partial regression
beta_hat2 <- solve(t(X22) %*% X22) %*% t(X22) %*% y11

#Beta hat 2 from partial regression
beta_hat1 <- solve(t(X1) %*% X1) %*% t(X1) %*% y - solve(t(X1) %*% X1) %*% t(X1) %*% X2 %*% beta_hat2


beta_hat
beta_hat1
beta_hat2
```

As you can see, the values of $\hat{\beta}_0...\hat{\beta}_5$ from full regression match the values of $\hat{\beta}_0...\hat{\beta}_3$ and $\hat{\beta}_4, \hat{\beta}_5$ found using partial regression.

### Question j

Lagrange
```{r}
ones <- rep(1, nrow(a))
X <- as.matrix(cbind(ones,x1, x2, x3, x4))
C <- as.matrix(rbind(c(1,2,-3,5,-1), c(2,-1,1,3,0)))
beta_hat <- solve(t(X) %*% X) %*% t(X) %*% y
gamma <- as.matrix(rbind(5,10))

#Beta hat using the constrained model
beta_hat_c <- beta_hat - solve(t(X) %*% X) %*% t(C) %*% solve(C %*% solve(t(X) %*% X) %*% t(C)) %*% (C %*% beta_hat - gamma)
beta_hat_c
```


Canonical 
```{r}
C1 <- as.matrix(rbind(c(1,2), c(2,-1)))
C2 <- as.matrix(rbind(c(-3,5,-1), c(1,3,0)))
X1 <- as.matrix(cbind(ones,x1))
X2 <- as.matrix(cbind(x2, x3, x4))


Yr <- y - X1 %*% solve(C1) %*% gamma
X2r <- X2 - X1 %*% solve(C1) %*% C2

#Beta hat 1 and beta hat 2 using the constrained model
beta_hat_2c <- solve(t(X2r) %*% X2r) %*% t(X2r) %*% Yr
beta_hat_1c <- solve(C1) %*% (gamma - C2 %*% beta_hat_2c)

beta_hat_1c
beta_hat_2c
```

As you can see, the values of $\hat{\beta}_{c0}...\hat{\beta}_{c4}$ obtained using Lagrange multipliers are the same as the values of $\hat{\beta}_{c0}, \hat{\beta}_{c1}$ and $\hat{\beta}_{c2}...\hat{\beta}_{c4}$ obtained using the canonical form of the model. 



